Type theories as quotient inductive-recursive types

### 2025-03-07 tex/main.tex
mathematical typo
Minor wording change

To retain the local confluence, we turn $\sigma \uparrow A$ into a simultaneously defined function:
\begin{alignat}{3}
  \blank\uparrow \blank & : \implicit{\Gamma, \Delta}\; (\sigma : \Sub\,\Gamma,\Delta) (A : \Ty\,\Delta) \to \Sub\,(\Gamma, \subTy{A}{\sigma})\, (\Delta, A) \label{eq:type-of-lifting} \\
\idS                \uparrow A & \reduce \idS \nonumber \\
(\sigma ; \tau)     \uparrow A & \reduce (\sigma \uparrow \subTy{\tau}{A} ) ; (\tau \uparrow A) \nonumber \\
\pi_1(\sigma, t)    \uparrow A & \reduce \sigma \uparrow A \nonumber \\
\pi_1(\sigma; \tau) \uparrow A & \reduce (\sigma \uparrow \subTy{\pi_1 \tau}{A}) ; (\pi_1 \tau \uparrow A)\nonumber \\
\sigma              \uparrow A & \reduce \sigma^+ , \quad \text{otherwise} \nonumber
\end{alignat}
where $\sigma^+ \defeq (\pi_1\idS ; \sigma), \pi_2\idS$.
Then, we add another clause for type substitution
\begin{equation*}
  \subTy{(\Pi\,A\,B)}{\sigma} \reduce \Pi\,(\subTy{A}{\sigma})\,(\subTy{B}{\sigma\uparrow A}) %\label{eq:def-type-subst-9}
\end{equation*}
and an equality constructor for term substitution without transport:
\begin{alignat}{5}
  \mathsf{[]\lambda} :{} & \sub{(\lambda\,t)}{\sigma} && =^{\Tm\,\Gamma\,(\subTy{(\Pi\,A\,B)}{\sigma})} && \lambda\,(\sub{t}{\sigma \uparrow A}) \nonumber
\end{alignat}

@@

\subparagraph*{Implementing overlapping patterns via rewrite rules}
%The mechanism of defining a function by overlapping patterns is only implemented in an outdated experimental fork of \Agda.
To emulate a function by overlapping patterns, we implement each function clause as a rewriting rule, together with another rule to identify the clause as the reflexivity proof $\mathsf{refl}$ (which ensures that the named function clause is merely synonymous with \lstinline|refl|).
For example, the first few clauses for type substitution (written $[\sigma]_{\Ty}\,A$ rather than $\subTy{A}{\sigma}$ in the formalisation) are defined as
\begin{lstlisting}
postulate
  [_]_      : Sub Γ Δ → Ty Δ → Ty Γ
  [id]      : [ idS ]    A ≡ A
  {-# REWRITE [id]      #-}
  [id]=refl : [id] {A = A} ≡ refl
  {-# REWRITE [id]=refl #-}
\end{lstlisting}
We also turn on the option \texttt{-{}-local-confluence-check} to check local confluence globally.
%
Overlapping patterns is indeed one of the use cases of user-defined rewrite rules discussed by Cockx~\cite{Cockx2020}.
The caveat, though, is that the termination checking is not imposed, so we need to check termination informally.

>>>>

To retain the local confluence, we turn $\sigma \uparrow A$ into a simultaneously defined function:
\begin{alignat}{3}
  \blank\uparrow \blank & : \implicit{\Gamma, \Delta}\; (\sigma : \Sub\,\Gamma\,\Delta) (A : \Ty\,\Delta) \to \Sub\,(\Gamma, \subTy{A}{\sigma})\, (\Delta, A) \label{eq:type-of-lifting} \\
\idS                \uparrow A & \reduce \idS \nonumber \\
(\sigma ; \tau)     \uparrow A & \reduce (\sigma \uparrow \subTy{\tau}{A} ) ; (\tau \uparrow A) \nonumber \\
\pi_1(\sigma, t)    \uparrow A & \reduce \sigma \uparrow A \nonumber \\
\pi_1(\sigma; \tau) \uparrow A & \reduce (\sigma \uparrow \subTy{\pi_1 \tau}{A}) ; (\pi_1 \tau \uparrow A)\nonumber \\
\sigma              \uparrow A & \reduce \sigma^+ , \quad \text{otherwise} \nonumber
\end{alignat}
where $\sigma^+ \defeq (\pi_1\idS ; \sigma), \pi_2\idS$.
Then, we add another clause for type substitution
\begin{equation*}
  \subTy{(\Pi\,A\,B)}{\sigma} \reduce \Pi\,(\subTy{A}{\sigma})\,(\subTy{B}{\sigma\uparrow A}) %\label{eq:def-type-subst-9}
\end{equation*}
and an equality constructor for term substitution without transport:
\begin{alignat}{5}
  \mathsf{[]\lambda} :{} & \sub{(\lambda\,t)}{\sigma} && =^{\Tm\,\Gamma\,(\subTy{(\Pi\,A\,B)}{\sigma})} && \lambda\,(\sub{t}{\sigma \uparrow A}) \nonumber
\end{alignat}

@@

\subparagraph*{Implementing overlapping patterns via rewrite rules}
%The mechanism of defining a function by overlapping patterns is only implemented in an outdated experimental fork of \Agda.
To emulate a function by overlapping patterns, we implement each function clause as a rewriting rule, together with another rule to identify the clause as the reflexivity proof $\mathsf{refl}$ (which ensures that the named function clause is merely a synonym for \lstinline|refl|).
For example, the first few clauses for type substitution (written $[\sigma]_{\Ty}\,A$ rather than $\subTy{A}{\sigma}$ in the formalisation) are defined as
\begin{lstlisting}
postulate
  [_]_      : Sub Γ Δ → Ty Δ → Ty Γ
  [id]      : [ idS ]    A ≡ A
  {-# REWRITE [id]      #-}
  [id]=refl : [id] {A = A} ≡ refl
  {-# REWRITE [id]=refl #-}
\end{lstlisting}
We also turn on the option \texttt{-{}-local-confluence-check} globally to check local confluence.
%
Overlapping patterns is indeed one of the use cases of user-defined rewrite rules discussed by Cockx~\cite{Cockx2020}.
The caveat, though, is that the termination checking is not imposed, so we need to check termination informally.
